---
title: "Mozambique coastal analysis"
author: "Tim White"
date: "9/6/2019"
output: html_document
---


Mozambique requested an overview of fishing in their coastal waters. This file produces summary maps of fishing in the EEZ and surrounding waters.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message = FALSE, comment=FALSE, warning=FALSE}

library(fishwatchr)
library(tidyverse)
library(bigrquery)
library(ggsci) 
library(rnaturalearth)
library(sf)
library(raster)
library(viridis)
library(googledrive)

# Establish connection to BigQuery project
con <- DBI::dbConnect(bigrquery::bigquery(), project = "world-fishing-827", use_legacy_sql = FALSE)



setwd("~/analysis/africa/mozambique/code")

#bq_auth  # run this if bigquery account is not yet authorized 

```

Read in GFW logo and land shapefile

```{r}

setwd("~/analysis/africa/mozambique/code")

# GFW logo
gfw_logo <- png::readPNG("~/analysis/japan/data/GFW_logo_primary_White.png")
gfw_logo_rast <- grid::rasterGrob(gfw_logo, interpolate = T)

   
# note the ! in front of Line_type: select all lines NOT baselines   
eez_shp <- read_sf("/Users/timwhite/Downloads/World_EEZ_v10_20180221/") %>%
 filter(!Line_type %in% c("Straight Baseline","Archipelagic Baseline"))       

# 12 NM boundary that Charlie modified and sent to #gfw-analysis-cell on Nov 27 2019
NM12_shp <- read_sf("../Data/Mozambique_12nm_2/")

# only needed to write shp for the first time
#st_write(eez_shp, "eez200NM.shp")
#eez_shp = read_sf("/Users/timwhite/Desktop/eez_200NM/")

world = ne_countries(scale = "large", returnclass = "sf")
#land_shp <- sf::read_sf('../data/shapefiles/ne_10m_world_0_360_raw/ne_10m_world_0_360_raw.shp')



# read in license list from MOZ

licenses = read.csv("../data/MOZ_LicenseList_April19_Converted_normalized.csv", header = TRUE,stringsAsFactors = FALSE)

# Linear green color palette function
effort_pal <- colorRampPalette(c('#0C276C', '#3B9088', '#EEFF00', '#ffffff'),
                               interpolate = 'linear')


   
   
```





```{r}

world = ne_countries(scale = "medium", returnclass = "sf")

land_shp <- sf::read_sf('~/data-dev/analysis/npfc/data/shapefiles/ne_10m_world_0_360_raw/ne_10m_world_0_360_raw.shp')

indo_shp <- read_sf("/Users/timwhite/Dropbox/AIS/data/World_EEZ_v9_20161021_LR/", layer = "eez_lr") %>% # lower res file. too big to fit in Github so its in Dropbox
  filter(Pol_type == '200NM') %>%  # select the 200 nautical mile polygon layer 
  filter(Sovereign1 == "Indonesia")


eez_shp <- read_sf("/Users/timwhite/Downloads/World_EEZ_v10_20180221/") 

chile_sf = eez_sf %>%
  filter(Sovereign == "Chile")

nazca_sf <- sf::read_sf("../data/WDPA_Jul2020_protected_area_555624169-shapefile/WDPA_Jul2020_protected_area_555624169-shapefile-polygons.shp")


nazca_map = ggplot() +
   geom_sf(data = chile_sf,
    color = 'purple',
    alpha = 0.3,
    fill = "purple",
    size = 0.3) + 
   geom_sf(data = world,
        fill = 'black',
        color = 'black',
        size = 0.1) +
  geom_sf(data = nazca_sf,
    color = 'purple',
    fill = "purple",
    size = 0.5) +
  #coord_equal() + # ensures that one unit on the x-axis is the same length as one unit on the y-axis
  theme_void()   # set empty theme

ggplot2::ggsave("../revision_figs/nazca_map.png",
                nazca_map,
                dpi = 600,
                width = 7,
                height = 5)

world_map = ggplot() +
   geom_sf(data = world,
        fill = 'black',
        color = 'black',
        size = 0.1) +
 
  theme_void()   # set empty theme

ggplot2::ggsave("../revision_figs/world_map.png",
                world_map,
                dpi = 600,
                width = 7,
                height = 5)


```








```{r}
ross_sf <- sf::read_sf("../data/WDPA_Apr2020_protected_area_555624810-shapefile/WDPA_Apr2020_protected_area_555624810-shapefile-polygons.shp")

nazca_sf <- sf::read_sf("../data/WDPA_Jul2020_protected_area_555624169-shapefile/WDPA_Jul2020_protected_area_555624169-shapefile-polygons.shp")




#get shapefile of MPA - TAKES THE MOST TIME - about 2 mins 16 seconds bruh
mpa <- readOGR(dsn = "/Users/tiffanyong/Desktop/MPA_shapefiles/", layer = "WDPA_June2018_marine-shapefile-polygons")



#TODO: set mpa name/creation date/mpant ------
###########VISUALIZING THE MPA#############
creation_date = '2017-12-01'
mpa_name = 'Ross Sea Region Marine Protected Area'
curr_mpa = mpa[mpa@data$NAME == 'Ross Sea Region Marine Protected Area', ]
curr_mpa = curr_mpa[1:3, ]
curr_mpa = spTransform(curr_mpa, CRS("+proj=longlat +datum=WGS84 +lon_wrap=180"))
#View(curr_mpa)
mpant = '9047'

#vizualize the MPA
leaflet() %>%
  addTiles() %>%
  addPolygons(data = ross_sf, color ="green")
```


```{r}

mpa_monthly_totals %>%
  filter(mpa_monthly_totals)
str(mpa_monthly_totals[[1]])
  

filelist <- mapply(cbind, mpa_monthly_totals, "SampleID"=ID, SIMPLIFY=F)


nameVec = names(mpa_monthly_totals)

for(i in 1:length(nameVec)){
  testmpa_monthly_totals[[nameVec[i]]]$mpa = nameVec[i]
  return(mpa_monthly_totals[[i]][1:12,])
}


mean(mpa_monthly_totals[[1]][1:12,"fishing_effort"], na.rm = T)
mean(mpa_monthly_totals[[2]][1:12,"fishing_effort"], na.rm = T)

#PAPA
mean(mpa_monthly_totals[[3]][1:12,"fishing_effort"], na.rm = T)
mean(mpa_monthly_totals[[3]][13:24,"fishing_effort"], na.rm = T)

mean(mpa_monthly_totals[[4]][1:12,"fishing_effort"], na.rm = T)
mean(mpa_monthly_totals[[5]][1:12,"fishing_effort"], na.rm = T)

mean(c(0.07025463,4.694757,41.04339,1.968727))


mean()


test2 = do.call("rbind",mpa_monthly_totals)

first_12_only = test2 %>%
  filter(mpa %in% c("Nazca","Pacific Remote Islands","Papa","Pitcairn" )) %>%
  group_by(mpa) %>%
  mutate(month = row_number()) %>%
  filter(month < 13) %>%
#  mutate(mean = mean(fishing_effort, na.rm = T)) %>%
  summarise_at(vars(fishing_effort),funs(mean(.,na.rm = TRUE)))

overall_mean = mean(first_12_only$fishing_effort)

%>%
  mean(fishing_effort, )

mpa_monthly_totals

lapplyTest = lapply(mpa_monthly_totals, `[[`, 2, [1:12])

lapplyTest = lapply(mpa_monthly_totals, `[[`, 2)
lapplyTest

lapplyTest
mpa_monthly_totals[[1]][2]



```


The reviewers took issue with us using coastal areas as comparison for Nazca. Instead, I'm quering effort in the non-Nazca part of the Chilean EEZ (see map, there is a un unprotected slice of the isolated oceanic Chilean EEZ that we will use for comparison)

```{sql connection = con, output.var = "fishing_sanfelix"}
WITH

-- This subquery identifies good segments
  good_segments AS (
  SELECT
    seg_id
  FROM
    `gfw_research.pipe_v20190502_segs`
  WHERE
    good_seg
    AND positions > 10
    AND NOT overlapping_and_short),

--   This subquery fishing query gets gridded daily fishing effort of target vessels
  fishing AS (
    SELECT
    date,
    lat_bin,
    lon_bin,
    SUM(fishing_hours) as fishing_hours
    FROM(
      SELECT
      ssvid,
      FLOOR(lat / 0.1) * 0.1 as lat_bin,
      FLOOR(lon / 0.1) * 0.1 as lon_bin,
      nnet_score2,
      timestamp,
      EXTRACT(date FROM date) as date,
      hours,
      IF
    (ARRAY_LENGTH(regions.eez)> 0,
      regions.eez[ORDINAL(1)],
      NULL) AS eez,
      # Calculate fishing hours by evaluating neural net score
      IF(nnet_score2 > 0.5, hours, 0) as fishing_hours
      FROM
        # Query the _fishing table to reduce query size
        `gfw_research.pipe_v20190502_fishing`
      WHERE 
      date >= "2015-08-23"
      AND date <= "2017-08-23"
      AND "22756" in unNEST(regions.eez)   -- only query effort in part of Chilean EEZ
      AND seg_id IN (
        SELECT
          seg_id
        FROM
          good_segments)
      )
    GROUP BY date, lat_bin, lon_bin
    HAVING
    fishing_hours > 10)

#####################
# Return fishing data
SELECT *
FROM fishing

```

fishing_sanfelix

Since there are very few points, I'm manually reviewing to see which ones are inside the MPA are which ones aren't.


```{r}

fishing_sanfelix


#fishing_sanfelix_reviewed = fishing_sanfelix[c(2,5,7),] # pull only the points that are inside MPA

fishing_sanfelix_reviewed

fishing_sanfelix



```

NONE ARE INSIDE MPA. So it's literally all zeros for this chunk....


Now repeat for the artificially designated chunk of Chile's EEZ that extends West of the rest of it, at the insistence of reviewers.

```{sql connection = con, output.var = "fishing_fernandez"}
WITH

-- This subquery identifies good segments
  good_segments AS (
  SELECT
    seg_id
  FROM
    `gfw_research.pipe_v20190502_segs`
  WHERE
    good_seg
    AND positions > 10
    AND NOT overlapping_and_short),

--   This subquery fishing query gets gridded daily fishing effort of target vessels
  fishing AS (
    SELECT
    date,
    lat_bin,
    lon_bin,
    SUM(fishing_hours) as fishing_hours
    FROM(
      SELECT
      ssvid,
      FLOOR(lat / 0.1) * 0.1 as lat_bin,
      FLOOR(lon / 0.1) * 0.1 as lon_bin,
      nnet_score2,
      timestamp,
      EXTRACT(date FROM date) as date,
      hours,
      IF
    (ARRAY_LENGTH(regions.eez)> 0,
      regions.eez[ORDINAL(1)],
      NULL) AS eez,
      # Calculate fishing hours by evaluating neural net score
      IF(nnet_score2 > 0.5, hours, 0) as fishing_hours
      FROM
        # Query the _fishing table to reduce query size
        `gfw_research.pipe_v20190502_fishing`
      WHERE 
      date >= "2015-08-23"
      AND date <= "2017-08-23"
      AND "8465" in unNEST(regions.eez)   -- only query effort in part of Chilean EEZ
      AND lat > -37.2
      AND lat < -30.2
      AND lon < -75.5
      AND seg_id IN (
        SELECT
          seg_id
        FROM
          good_segments)
      )
    GROUP BY date, lat_bin, lon_bin
    HAVING
    fishing_hours > 10)

#####################
# Return fishing data
SELECT *
FROM fishing


```

Review to see which points are inside EEZ

```{r}
View(fishing_fernandez)


2017-06-29	

fishing_fernandez_reviewed = fishing_fernandez %>%
  filter(date < "2017-06-28") # pull only the points that are inside MPA ALL POINTS AFTER THIS DATE WERE OUTSIDE MPA since I queried a square bbox and the shape is diagonal


fishing_fernandez_monthly <- fishing_fernandez_reviewed %>%
    mutate(month = format(date, '%m'),
           year = format(date, '%Y'),
           mon_year = paste0(month,'-',year)) %>%
    group_by(mon_year) %>%
    summarize(total_fishing_hours = sum(fishing_hours, na.rm = TRUE)) %>%
    ungroup()
fishing_fernandez_monthly

#THIS IS THE DATA TO ADD TO EXISTING FIGURES
#unadjustest by area: Aug 2015 (month 1) = 102.5 hours
#                     Sept 2015 (month 2) - 20.2 hours


```



